---
title: "Midterm Project"
author: "Christopher Flores and Zack Chipperfield"
output:
  html_document:
    df_print: paged
  word_document: default
---

Library that we used for creating some of the histograms
```{r}
library(tidyverse)
```

Reading in the data set from Github:
```{r, read_data}
data_url <- 'https://raw.githubusercontent.com/florescss14/STAT1341_Midterm_Project/main/Bundesliga.csv'
bund <- readr::read_csv(data_url, col_names = TRUE)

data_url <- 'https://raw.githubusercontent.com/florescss14/STAT1341_Midterm_Project/main/La_Liga.csv'
liga <- readr::read_csv(data_url, col_names = TRUE)

data_url <- 'https://raw.githubusercontent.com/florescss14/STAT1341_Midterm_Project/main/Ligue_1.csv'
ligue <- readr::read_csv(data_url, col_names = TRUE)

data_url <- 'https://raw.githubusercontent.com/florescss14/STAT1341_Midterm_Project/main/Premier_League.csv'
epl <- readr::read_csv(data_url, col_names = TRUE)

data_url <- 'https://raw.githubusercontent.com/florescss14/STAT1341_Midterm_Project/main/Serie_A.csv'
serie <- readr::read_csv(data_url, col_names = TRUE)

```
As some leagues have a different number of matches played we want to normalize it to number of points scored per match in each league.
```{r}
bund$Pts.MP <- bund$Pts / bund$MP
liga$Pts.MP <- liga$Pts / liga$MP
ligue$Pts.MP <- ligue$Pts / ligue$MP
epl$Pts.MP <- epl$Pts / epl$MP
serie$Pts.MP <- serie$Pts / serie$MP


all <- rbind(bund, liga, ligue, epl, serie)
```

Figure 1: Histograms of Total Completion Percentage of All Passes for Clubs from all five major leagues from 2017/18 to 2020/21 season
```{r}
ggplot() +
geom_histogram(mapping = aes(x = bund$Total_Cmp_Pct), bins = 8) 

ggplot() +
geom_histogram(mapping = aes(x = liga$Total_Cmp_Pct), bins = 8) 

ggplot() +
  geom_histogram(mapping = aes(x = ligue$Total_Cmp_Pct), bins = 8)
ggplot() +
  geom_histogram(mapping = aes(x = epl$Total_Cmp_Pct), bins = 8)
ggplot() +
  geom_histogram(mapping = aes(x = serie$Total_Cmp_Pct), bins = 8)
```
Figure 2. Mean and Standard Deviation of Total Pass Completion Percentage of Clubs in Major European Leagues

```{r}
mean(bund$Total_Cmp_Pct)
sd(bund$Total_Cmp_Pct)
mean(liga$Total_Cmp_Pct)
sd(liga$Total_Cmp_Pct)
mean(ligue$Total_Cmp_Pct)
sd(ligue$Total_Cmp_Pct)
mean(epl$Total_Cmp_Pct)
sd(epl$Total_Cmp_Pct) 
mean(serie$Total_Cmp_Pct)
sd(serie$Total_Cmp_Pct)
```
Figure 3: Correlation Values between Total Pass Completion Percentage and Points per Match for Clubs in Major European Leagues from 2017/18 to 2020/21 seasons


```{r}
cor(bund$Total_Cmp_Pct, bund$Pts.MP)
cor(liga$Total_Cmp_Pct, liga$Pts.MP)
cor(ligue$Total_Cmp_Pct, ligue$Pts.MP)
cor(epl$Total_Cmp_Pct, epl$Pts.MP)
cor(serie$Total_Cmp_Pct, serie$Pts.MP)
```

Looking at the correlations for short, medium and long completion percentages versus points per match.
```{r}
cor(bund$Short_Cmp_Pct, bund$Pts.MP)
cor(liga$Short_Cmp_Pct, liga$Pts.MP)
cor(ligue$Short_Cmp_Pct, ligue$Pts.MP)
cor(epl$Short_Cmp_Pct, epl$Pts.MP)
cor(serie$Short_Cmp_Pct, serie$Pts.MP)
```

```{r}
cor(bund$Medium_Cmp_Pct, bund$Pts.MP)
cor(liga$Medium_Cmp_Pct, liga$Pts.MP)
cor(ligue$Medium_Cmp_Pct, ligue$Pts.MP)
cor(epl$Medium_Cmp_Pct, epl$Pts.MP)
cor(serie$Medium_Cmp_Pct, serie$Pts.MP)
```

```{r}
cor(bund$Long_Cmp_Pct, bund$Pts.MP)
cor(liga$Long_Cmp_Pct, liga$Pts.MP)
cor(ligue$Long_Cmp_Pct, ligue$Pts.MP)
cor(epl$Long_Cmp_Pct, epl$Pts.MP)
cor(serie$Long_Cmp_Pct, serie$Pts.MP)
```

Figure 4. All Five Leagues Summary off Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression Points per Match, from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=all)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = all)
step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(step_model)
```

Calculating predicted Points per match and the error
```{r}
all$pred_Pts.MP <- predict(step_model, all)
all$error <- all$Pts.MP - all$pred_Pts.MP
```

Figure 5: Plot of Actual Points per Match vs. Predicted Points per Match (using the model in Figure 4) of All Clubs in Major Five Leagues from 2017/18 to 2020/21 season
```{r}
plot(all$Pts.MP, all$pred_Pts.MP)
```

Figure 6: Res Plot of Overall Regression
```{r}
plot(all$pred_Pts.MP, all$error)
```

Figure 7:  Bundesliga Summary of Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression on Points per Match, accounting for all 5 major leagues from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=bund)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = bund)
bund_step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(bund_step_model)

bund$pred_Pts.MP <- predict(bund_step_model, bund)
bund$error <- bund$Pts.MP - bund$pred_Pts.MP
```

Figure 8: English Premier League Summary of Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression on Points per Match, from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=epl)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = epl)
epl_step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(epl_step_model)

epl$pred_Pts.MP <- predict(epl_step_model, epl)
epl$error <- epl$Pts.MP - epl$pred_Pts.MP
```

Figure 9: La Liga Summary of Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression on Points per Match, from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=liga)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = liga)
liga_step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(liga_step_model)

liga$pred_Pts.MP <- predict(liga_step_model, liga)
liga$error <- liga$Pts.MP - liga$pred_Pts.MP
```

Figure 10: Ligue 1 Summary of Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression Points per Match, accounting for all 5 major leagues from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=ligue)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = ligue)
ligue_step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(ligue_step_model)

ligue$pred_Pts.MP <- predict(ligue_step_model, ligue)
ligue$error <- ligue$Pts.MP - ligue$pred_Pts.MP
```

Figure 11: Serie A Summary of Final Model of Coefficient Estimates, Standard Errors, Test Statistics, P-Values, and Significance for Linear Regression Points per Match, accounting for all 5 major leagues from 2017/18 to 2020/21 seasons
```{r}
intercept_only <- lm(Pts.MP ~ 1, data=serie)
all_predictors <- lm(Pts.MP ~ Total_Cmp_Pct  + Ent_Final_Third + Ent_Opposing_PA + Cross_into_PA + PrgDist + Short_Cmp_Pct + Medium_Cmp_Pct + Long_Cmp_Pct + Prog_10_plus_yards + KP, data = serie)
serie_step_model <- step(intercept_only, direction = "both", scope=formula(all_predictors), trace=0)
summary(serie_step_model)

serie$pred_Pts.MP <- predict(serie_step_model, serie)
serie$error <- serie$Pts.MP - serie$pred_Pts.MP
```

Figure 12: Plot of Actual Points per Match vs. Predicted Points per Match (using the models indicated in Figure 7-11) of All Clubs in Major Five Leagues from 2017/18 to 2020/21 season
```{r}
par(mfrow=c(2,3))
plot(bund$Pts.MP, bund$pred_Pts.MP)
plot(epl$Pts.MP, epl$pred_Pts.MP)
plot(liga$Pts.MP, liga$pred_Pts.MP)
plot(ligue$Pts.MP, ligue$pred_Pts.MP)
plot(serie$Pts.MP, serie$pred_Pts.MP)
```

Figure 13: Res Plot of all 5 individual leagues
```{r}
par(mfrow=c(2,3))
plot(bund$pred_Pts.MP, bund$error)
plot(epl$pred_Pts.MP, epl$error)
plot(liga$pred_Pts.MP, liga$error)
plot(ligue$pred_Pts.MP, ligue$error)
plot(serie$pred_Pts.MP, serie$error)
```










